---
title: "Colormaps"
output:
  html_document:
    toc: true
    toc_depth: 2
---

[Colorbrewer][1] and [viridis][2] have been the two main resources for this page.

```{r colormaps_colorbrewer_colorblind, fig.height=8}
library(RColorBrewer)
display.brewer.all(colorblindFriendly = TRUE)
```


# Continuous (sequential)

Please see the [viridis: Default Color Maps from 'matplotlib' CRAN package][2].
Thre is a [great vignette][3] by the package authors.

```{r colormaps_ggplot_virdis_fill}
library(ggplot2)
library(viridis)

ggplot(data.frame(x = rnorm(10000), y = rnorm(10000)), aes(x = x, y = y)) +
    geom_hex() +
    coord_fixed() +
    scale_fill_viridis() + # The virdis color scale
    theme_bw()
```

## Decile color palette

```{r colormap_virdis_decile}
library(scales)
show_col(viridis_pal(option = 'virdis')(10))
```

```{r colormap_magma_decile}
show_col(viridis_pal(option = 'magma')(10))
```

# Divergent

Here's a function to plot colors in a series of rectangles (written by Vicki)

```{r}
pal <- function(col, border = "light gray", ...) {
    n <- length(col)
    plot(0, 0,
         type = "n",
         xlim = c(0, 1),
         ylim = c(0, 1),
         axes = FALSE,
         xlab = "", ylab = "", ...)
    rect(0:(n - 1) / n,
         0,
         1:n/n,
         1,
         col = col,
         border = border)
}
```

HEX
```
#b2182b
#ef8a62
#fddbc7
#d1e5f0
#67a9cf
#2166ac
```

```{r colormap_divergent}
colours <- c(
    '#b2182b',
    '#ef8a62',
    '#fddbc7',
    '#d1e5f0',
    '#67a9cf',
    '#2166ac'
)
pal(colours)
```



# Categorical (qualiative)

Here is a function to plot colors in a pie chart (written by Vicki)

```{r}
wheel <- function(col, radius = 1, ...) {
    pie(rep(1, length(col)), col = col, radius = radius, ...)
}
```

## Colorblind and Print friendly

[Colorbrewer][1] only has 1 color scheme that is both colorblind and print friendly.
It only allows 4 classes.


Hex:
```
#a6cee3
#1f78b4
#b2df8a
#33a02c
```

```{r colormap_categorical_print_cb}
cbPalette <- c(
    '#a6cee3',
    '#1f78b4',
    '#b2df8a',
    '#33a02c'
)

wheel(cbPalette)
```



Here we added a grey color to be used as a `NA` value in plots.

```{r colormap_categorical_print_cb_bgw}
cbPalette <- c(
    '#a6cee3',
    '#1f78b4',
    '#b2df8a',
    '#33a02c',
    "#999999",
    '#000000',
    '#ffffff'
)

wheel(cbPalette)
```



RGB:
```
166,206,227
31,120,180
178,223,138
51,160,44
```

CMTK
```
35,7,0,0
90,30,0,0
30,0,45,0
80,0,100,0
```

## Colorblind friendly

The colorbline friendly pallete is the same as the colorblind and print friendly color scheme.

## Print friendly

Removing the colorblind friendly limitation, gives us a maximum of 9 qualitative colors.
There is only 1 color scheme for this many classes.

HEX
```
#e41a1c
#377eb8
#4daf4a
#984ea3
#ff7f00
#ffff33
#a65628
#f781bf
#999999
```

```{r colormap_categorical_print}
cbPalette <- c(
    '#e41a1c',
    '#377eb8',
    '#4daf4a',
    '#984ea3',
    '#ff7f00',
    '#ffff33',
    '#a65628',
    '#f781bf',
    '#999999'
)

wheel(cbPalette)
```


RGB
```
228,26,28
55,126,184
77,175,74
152,78,163
255,127,0
255,255,51
166,86,40
247,129,191
153,153,153
```

CMYK
```
10,90,80,0
80,30,0,0
70,0,80,0
40,65,0,0
0,50,100,0
0,0,80,0
35,60,80,0
0,50,0,0
0,0,0,40
```


[1]: http://colorbrewer2.org
[2]: https://cran.r-project.org/web/packages/viridis/index.html
[3]: https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
